基于卷积神经网络的视觉目标跟踪算法研究

### 2021-03-22 博士/毕业论文/ucasthesis/Tex/Frontmatter.tex
改英文

Visual object tracking is one of the most important and challenging research topics in the field of computer vision, and it has potential and broad application prospects in the fields of intelligent surveillance and autonomous driving. The goal of visual object tracking is to estimate the motion state of the target in each frame of the image sequences. Visual object tracking is the middle level of the computer vision field, which provides a basis for the understanding of target behavior, so it has very important theoretical research value. At the same time, it has a wide range of practical applications, including video surveillance, traffic flow monitoring, video compression, and human-computer interaction.
However, there are many theoretical and technical problems to be solved in the field of object tracking, such as motion blur, illumination changes, deformation of non-rigid targets, target rotation and occlusion caused by changes in perspective. In recent years, the breakthrough of deep learning has brought the possibility to solve a series of problems in object tracking.
Deep learning is a machine learning method based on artificial neural networks. In the past ten years, deep learning technology has been rapidly developed and has been successfully applied in computer vision, speech recognition, natural language processing, audio recognition, social network filtering, machine translation, bioinformatics, drug design and other fields. How to use deep learning methods, especially deep convolutional neural networks to solve the complex problems encountered in the tracking process, has great research value and research space.

This paper uses the powerful representation ability of convolutional neural network to improve the feature representation of video object tracking algorithm, appearance model modeling, motion model modeling, model adaptation, etc., and effectively improve the performance of the algorithm. At the same time, this article applies the adversarial information to the video object tracking task to study the robustness of the video object tracking algorithm relative to the adversarial samples. The main work and contributions of this paper are summarized as follows:

\begin{itemize}
\item{\textbf{We propose a video object tracking algorithm guided by semantic information.} The proposed algorithm uses the convolutional neural network to obtain the \textbf{semantic information} of the target, which is used to constrain the training process of the tracker, thereby improving the tracking effect. Specifically, we first propose an instance-guided correlation filter, which uses a convolutional neural network to learn the semantic segmentation template at the instance level of the image, thereby constraining the learning of the correlation filter. Secondly, in view of the complementarity between the semantic segmentation results of offline training and the correlation filtering results of online learning, we propose a self-correction mechanism for tracking results, using the segmentation results to correct the correlation filtering results. We have verified the effectiveness of these innovations in video object tracking applications on multiple challenging video object tracking databases.}

\item{\textbf{We propose a visual object tracking algorithm with enhanced spatial information.} The proposed algorithm mainly optimizes the video object tracking algorithm based on the Siamese network in the aspect of motion model. Specifically, we introduce richer \textbf{spatial information} for the Siamese tracking network, that is, always perceive the position information of the object in the entire image plane, which can make up for the shortcomings of the limited target search range in the traditional local search mechanism to reduce accumulated errors and improve robustness. In order to further reduce the interference of similar objects, we propose an end-to-end training trajectory prediction module that can use the historical trajectory information of the object and the apparent information of the current frame to predict the possibility of the target appearing in each spatial position of the current frame. We have verified the effectiveness of these innovations on multiple video tracking standard evaluation libraries, and have greatly improved the accuracy and robustness of the tracking algorithm.}

\item{\textbf{We propose a visual object tracking algorithm enhanced by time information.} The proposed algorithm mainly optimizes the video object tracking algorithm based on Siamese network in feature extraction. First, we start from the problem of insufficient robustness of the online video tracking algorithm based on the Siamese network, and introduce \textbf{time information} into online video object tracking. Through the aggregation of the target's appearance information from adjacent frames, the target's apparent features are enriched. It makes up for the shortcomings of the visual object tracking algorithm based on the Siamese network that is limited to extracting the target appearance from a single frame, and the lack of ability to express the target appearance, thereby improving the tracking effect and achieving robust tracking. On the basis of end-to-end time aggregation, we introduced the adversarial dropout module, and through end-to-end training on large-scale datasets, so that the Siamese network tracker has better performance in the case of poor appearance of the target due to motion blur, etc., thereby further improving the robustness of tracking. We have conducted algorithm comparison experiments and component analysis experiments on the currently popular video object tracking evaluation library to verify the effectiveness of the algorithm improvement.}

\item{\textbf{We propose an adaptive information-enhanced video object tracking algorithm} The proposed algorithm mainly optimizes the video object tracking algorithm based on the Siamese network in the aspect of model adaptation. We introduced \textbf{adaptive information} to the Siamese tracking network, which improves the tracking performance of the Siamese network tracker for specific targets by slightly perturbing the pixels of the template image. The adaptive information is obtained by performing gradient back propagation calculation on the template image, and can be easily added to the existing Siamese network tracker in a plug-and-play manner, without modifying the parameters of the network model. In online tracking, only a few times of gradient propagation and template image pixel value update in the first frame can achieve real-time object tracking. We also verified the effectiveness of the algorithm on multiple video object tracking evaluation libraries, and achieved good results in accuracy and real-time performance.}

\item{\textbf{We apply the adversarial information to the video object tracking algorithm based on the Siamese network.} The proposed algorithm applies \textbf{adversarial information} to the Siamese tracking network to study the robustness of the video object tracking algorithm relative to adversarial samples. Specifically, we generate video-independent universal perturbations for the video object tracking algorithm based on the Siamese network, which makes the tracker behave incorrectly. The proposed adversarial perturbation information is obtained through offline large-scale video object tracking dataset training, which can effectively attack any video while occupying very few computing resources. We verified the effectiveness of the proposed adversarial information on multiple video object tracking standard evaluation libraries, and at the same time verified its transferability in different backbone networks and different tracking frameworks.}
\end{itemize}

\KEYWORDS{Visual Object Tracking, Deep Learning, Conlvolutional Neural Network, Siamese Network, Correlation Filter}% 英文关键词
%---------------------------------------------------------------------------%

>>>>

Visual object tracking is one of the most important and challenging research topics in the field of computer vision, and it has potential and broad application prospects in the fields of intelligent surveillance and autonomous driving. The goal of visual object tracking is to estimate the motion state of the target in each frame of the image sequences. Visual object tracking is the middle level of the computer vision field, which provides a basis for activity understanding, so it has very important theoretical research value. At the same time, it has a wide range of practical applications, including video surveillance, traffic flow monitoring, video compression, and human-computer interaction.
However, there are many theoretical and technical problems to be solved in the field of object tracking, such as motion blur, illumination changes, deformation of non-rigid targets, target rotation and occlusion caused by changes in perspective. In recent years, the breakthrough of deep learning has brought the possibility to solve a series of problems in object tracking.
Deep learning is a machine learning method based on artificial neural networks. In the past ten years, deep learning technology has been rapidly developed and has been successfully applied in computer vision, speech recognition, natural language processing, audio recognition, social network filtering, machine translation, bioinformatics, drug design and other fields. How to use deep learning methods, especially deep convolutional neural networks to solve the complex problems encountered in the tracking process, has great research value and research space.

In this thesis, we utilize the powerful representation ability of convolutional neural networks to improve the visual object tracking algorithm from different aspects including feature representation, appearance model, motion model and model adaptation. Besides, we apply the adversarial information to the visual object tracking task to study the robustness of the visual object tracking algorithm with respect to the adversarial samples. The main work and contributions of this thesis are summarized as follows:

\begin{itemize}
\item{\textbf{We propose a semantic information guided visual object tracking algorithm.} The proposed algorithm uses the convolutional neural network to obtain the \textbf{semantic information} of the target, which is used to constrain the training process of the tracker to improve the tracking performance. Specifically, we first propose an instance-guided correlation filter, which uses a convolutional neural network to learn the semantic segmentation template of the target, thereby constraining the learning of the correlation filter. Secondly, in view of the complementarity between the semantic segmentation results and the correlation filter results, we propose a self-correction mechanism for tracking results, utilizing the segmentation results to correct the correlation filter results. Comprehensive experiments are performed on challenging visual object tracking datasets to validate the effectiveness of these innovations.}

\item{\textbf{We propose a spatial information enhanced visual object tracking algorithm.} The proposed algorithm mainly optimizes the motion model of the siamese object tracking algorithm. Specifically, we introduce richer \textbf{spatial information} for the siamese tracking network, that is, always perceiving the position information of the target in the entire image plane, which can make up for the shortcomings of the limited target search range in the traditional local search mechanism to reduce accumulated errors and improve robustness. In order to further reduce the interference of similar targets, we propose an end-to-end trajectory prediction module that can use the historical trajectory information of the target and the apparent information of the current frame to predict the possibility of the target appearing in each spatial position. Comprehensive experiments are performed on challenging visual object tracking datasets to validate the effectiveness of these innovations.}

\item{\textbf{We propose a temporal information enhanced visual object tracking algorithm.} The proposed algorithm mainly optimizes the feature representation of the siamese object tracking algorithm. First, we start from the problem of insufficient robustness of the online siamese tracking algorithm, and introduce \textbf{temporal information} into online visual object tracking. Through the aggregation of the target's appearance information from adjacent frames, the target's apparent features are enriched. It makes up for the shortcomings of the siamese object tracking algorithm's insufficient ability to represent the target appearance, thereby achieving robust tracking. On the basis of end-to-end temporal information aggregation, we further introduce the adversarial dropout module to improve the performance of the tracker in the case of the poor appearance of the target. Comprehensive experiments are performed on challenging visual object tracking datasets to validate the effectiveness of these innovations.}

\item{\textbf{We propose an adaptive information enhanced visual object tracking algorithm.} The proposed algorithm mainly optimizes the model adaptation of the siamese object tracking algorithm. We introduce \textbf{adaptive information} to the siamese tracking network, which improves the tracking performance of the siamese tracker for specific targets by slightly perturbing the pixels of the template image. The adaptive information is obtained by performing gradient back propagation on the template image, and can be easily added to the existing siamese network tracker in a plug-and-play manner, without modifying the parameters of the network model. Model adaptation can be achieved by only performing a few steps of gradient propagation and template image pixel value update in the first frame. Experimental results of challenging visual object tracking datasets demonstrate that our algorithm performs competitively against existing state-of-the-art methods and achieves high robustness and efficiency.}

\item{\textbf{We apply the adversarial information to the visual object tracking algorithm.} The proposed algorithm applies \textbf{adversarial information} to the siamese tracking network to study the robustness of the visual object tracking algorithm with respect to adversarial samples. Specifically, we generate video-independent universal perturbations for the siamese object tracking algorithm, which makes the tracker behave incorrectly. The proposed adversarial perturbation information is obtained through offline large-scale visual object tracking dataset training, which can effectively attack any video while occupying very few computing resources. Experimental results of challenging visual object tracking datasets demonstrate the effectiveness of the proposed adversarial information.}
\end{itemize}

\KEYWORDS{Visual Object Tracking, Deep Learning, Convolutional Neural Network, Siamese Network, Correlation Filter}% 英文关键词
%---------------------------------------------------------------------------%
